---
title: "In-class_Ex02"
author: "Qu JunJie"
date: "02 Sep 2024"
date-modified: "last-modified"
execute: 
  eval: true
  echo: true
  message: false
  freeze: true
format: 
  html:
    code-fold: true
---

```{r}
pacman::p_load(sf, raster, spatstat, tmap, tidyverse)
```

# Issue 1: Installing maptools

maptools is retired and bnary is removed from CRAN. However, we can download from Posit Public Package Manger snapshots by using the code chunk below.

```{r}
#| eval: FALSE
install.packages("maptools", repos = "https://packagemanager.posit.co/cran/2023-10-13")
```

After the installation is completed, it is important to edit the code chunk as show above in order to avoid

# Issue 2: Creating coastal outline

in sf package, there are two functions allow us to combine multuple simple features into

Working with st_union()

The code chunk below, st_union() is used to derive the coastal outline sf tibble data.frame.

```{r}
# sg_sf <- mpsz_sf %>%
#   st_union()
```

sg_sf will look similar to figure below.
